---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## load required packages
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Loading and preprocessing the data
## =====================================
* capture source file variable 
* capture url and download the zip file, if not already exists.
* unzip the source data and get the contents of the zip file.
* create data frame with read to csv
```{r FileAccess}
  file <- "./data/activity_monitoring_data.zip"
  if (!file.exists(file)){
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"  
    download.file(url,file)  
    }
  unzip(file, exdir="./data")
  data<- read.csv("./data/activity.csv")
 ```

## What is mean total number of steps taken per day?
## ==================================================
### create summary dataset
```{r }
  data1<- with(data, aggregate(steps, by=list(date), sum, na.rm =TRUE)) 
  names(data1)<- c("date", "total.steps")
```

### Make a histogram of the total number of steps taken each day ## Calculate and report the mean and median of the total number of steps taken per day
```{r}
  barplot(data1$total.steps, names.arg= data1$date, col= "blue", border= "red", xlab= "date", ylab= "total steps", main= "total steps per day")
  summary(data1$total.steps) 
  abline(h= summary(data1$total.steps)[["Mean"]], col= "red", lty=5, lwd=3) 
  abline(h= summary(data1$total.steps)[["Median"]], col= "green", lty=5, lwd=3) 
  legend("topleft", legend=c("Mean", "Median"), lty= 5, lwd=3, col= c("red", "green"))
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
